ST563 Final Project

### 2021-03-11 ST563-FP.Rmd
added in some code for some of the models. Still need to figure out what variables go where...

# Data
The data set used is the `Wine Quality` dataset from the UCI Machine Learning Repository. The data set is composed of 1600 observations of different variants of the Portuguese "Vinho Verde" red wine. For each wine, various features of the wine were measured, including a measurement of quality, given by an expert wine taster. Our objective it to use the different features to predict the quality of the wine.

@@

Fortunately, the `wine` data set that we worked with was very tidy; there was no missing values. We didn't standardize the data initially, but when training the support vector machine as well as the k-nearest neighbors models we centered and scaled the data.

@@

@@

The advantage of using a basic tree is that it is easy to understand and has a good interpretability, which is not something that translates over to ensemble methods like the random forest. Additionally, the basic classification or regression tree is not very computationally expensive, unlike the random forest, boosting, and the support vector machine. We used the `caret` and `rpart` packages in R to fit our trees.

@@

@@

@@

## Logistic Regression  
The logistic regression model uses the "logit" function $log(\frac{p}{1-p})$, which links the mean to the linear form of the regression model, $\boldsymbol{X\beta}$. Using it for binary classification, we round the fitted values either up or down to 1 or 0. The logistic regression function is defined as  
$$
\ln \left( \frac{p(x)}{1-p(x)} \right) = x'\beta, \qquad \text{where }p(x) = (1+e^{-x'\beta})
$$

@@

>>>>

# Data
The data set used is the `Wine Quality` dataset from the UCI Machine Learning Repository. The data set is composed of 1600 observations of different variants of the Portuguese "Vinho Verde" red wine. For each wine, various features of the wine were measured, including a measurement of quality, given by an expert wine taster. Our objective is to use the different features to predict the quality of the wine using different machine learning techniques.

@@

Fortunately, the `wine` dataset that we worked with was very tidy; there was no missing values. We didn't standardize the data initially, but when training the support vector machine as well as the k-nearest neighbors models we centered and scaled the data.

@@

```{r CV}
# Specify the cross-validation method (5-fold CV)
trctrl <- trainControl(method = "cv", number = 5)
```

@@

The advantage of using a basic tree is that it is easy to understand and has a good interpretability, which is not something that translates over to ensemble methods like the random forest. Additionally, the basic classification or regression tree is not very computationally expensive, unlike the random forest, boosting, and the support vector machine. We used the `caret` and `rpart` packages in R to fit our trees.
```{r tree, echo=FALSE}
#### REGRESSION ####
# Fit basic regression tree model
tree.reg <- train(form, data = trn, method = "rpart",
                  trControl = trctrl)

# Obtain RMSE and results
tree.reg.pred <- predict(tree.reg, newdata = tst)
tree.reg.rmse <- mean((tree.reg.pred - tst$quality)^2)

# Print out tree visualization
tree.reg.out <- rpart(form, data = trn, cp = 0.01798144)
rpart.plot(tree.reg.out, main="Regression Tree")

#### HOTEL 2 ####
# Fit classification tree model
tree.class <- train(form, data = trn, method = "rpart",
                    trControl = trctrl)

# Obtain confusion matrix and results for H2
tree.class.pred <- predict(tree.class, newdata = tst)
tree.class.CM <- confusionMatrix(tree.class.pred, tst$class)
tree.class.res <- c(tree.class.CM$overall[1], tree.class.CM$byClass[1:2])

# Print out tree visualization
tree.class.out <- rpart(form, data = trn, cp = 0.01798144)
rpart.plot(tree.class.out, main="Classification Tree")

# Print out confusion matrix
# list("Confusion Matrix for H1" = h1.tree.CM$table,
#      "Confusion Matrix for H2" = h2.tree.CM$table)
```

@@

```{r rf, echo=FALSE}
#### REGRESSION ####
# Formula for random forest
form.rf <-

# Fit the regression random forest model 
rf.reg <- train(form.rf, data = trn, method = "rf",
                trControl = trctrl, importance = TRUE)

# Obtain RMSE and results
rf.reg.pred <- predict(rf.reg, newdata = tst)
rf.reg.rmse <- mean((rf.reg.pred - tst$quality)^2)

#### CLASSIFICATION ####
# Fit the classification random forest model
rf.class <- train(form.rf, data = trn, method = "rf",
                trControl = trctrl, importance = TRUE)

# Obtain confusion matrix and results
rf.class.pred <- predict(rf.class, newdata = tst)
rf.class.CM <- confusionMatrix(rf.class.pred, tst$class)
rf.class.res <- c(rf.class.CM$overall[1], rf.class.CM$byClass[1:2])

# Print out confusion matrix
# list("Confusion Matrix for H1" = h1.rf.CM$table,
#      "Confusion Matrix for H2" = h2.rf.CM$table)
```

@@

```{r svm}
#### CLASSIFICATION ####
# Define formula for SVM model
form.svm <-

# Fit support vector machine for H1
svm.class <- train(form.svm, data = trn, method = "svmRadial", 
                   preProcess = c("center", "scale"),
                   trControl = trctrl, tuneLength = 10)

# Obtain confusion matrix and results for H1
svm.class.pred <- predict(svm.class, newdata = tst)
svm.class.CM <- confusionMatrix(svm.class.pred, tst$class)
svm.class.res <- c(svm.class.CM$overall[1], svm.class.CM$byClass[1:2])
```

@@

```{r knn}
#### REGRESSION ####
# Define formula for KNN model
form.knn <-

# Fit knn model for H1
knn.reg <- train(form.knn, data = trn, method = "knn",
                 trControl = trctrl, preProcess = c("center", "scale"))

# Obtain RMSE and results
knn.reg.pred <- predict(knn.reg, newdata = tst)
knn.reg.rmse <- mean((knn.reg.pred - tst$quality)^2)

#### CLASSIFICATION ####
# Fit knn model for H2
knn.class <- train(form.kn, data = trn, method = "knn",
                   trControl=trctrl, preProcess = c("center", "scale"))

# Obtain confusion matrix and results for H2
knn.class.pred <- predict(knn.class, newdata = tst)
knn.class.CM <- confusionMatrix(knn.class.pred, tst$class)
knn.class.res <- c(knn.class.CM$overall[1], knn.class.CM$byClass[1:2])
```

## Logistic Regression  
The logistic regression model uses the "logit" function, $log(\frac{p}{1-p})$, which links the mean to the linear form of the regression model, $\boldsymbol{X\beta}$. Using it for binary classification, we round the fitted values either up or down to 1 or 0. The logistic regression function is defined as  
$$
\ln \left( \frac{p(x)}{1-p(x)} \right) = x'\beta, \qquad \text{where }p(x) = (1+e^{-x'\beta})
$$

@@

# Results

# Conclusion
